### 2009-01-30 articles/2009/geoinformatics/article.tex
a few blanks removed (just a test)
git-svn-id: https://svn.osgeo.org/qgis/trunk@10042 c8812cc2-4d05-0410-92ff-de0c093fc19c
Added line wrapping since its easier to diff changes in shorter lines
git-svn-id: https://svn.osgeo.org/qgis/trunk@10055 c8812cc2-4d05-0410-92ff-de0c093fc19c
Edits from proof read through
git-svn-id: https://svn.osgeo.org/qgis/trunk@10056 c8812cc2-4d05-0410-92ff-de0c093fc19c

The QGIS project is the work of a group of dedicated developers,
translators, documenters, release helpers, bug reporters, and promoters. Their 
contributions are mainly on a voluntary basis, except in a few cases where 
contributors are able to work on QGIS as part of their daily work. QGIS 
is managed by the Project Steering Committee (PSC), a five member committee
providing technical guidance, community liason, release management, and
financial/marketing activities. The work of the QGIS project process is
spread between numerous people who each have a specific area of
resposibility, and ad hoc contributors.

These volunteers together with a large number of users make up the
world-wide QGIS comunity. Over time their efforts have resulted in a comprehensive,
valuable and useful code and documentation base which is free for everyone 
to use and improve on.

@@

With community platforms such as our website, wiki, forums and blog the QGIS project
provides latest news, release, usage, and development information. In most cases 
these community web sites permit user contributions after registering. 
The QGIS-user mailing, forum and Internet Relay Chat (IRC) provide a valuable interface 
with other users and for discussions of QGIS in general. In the spirit of open 
process and sharing knowledge, contacting developers directly instead of going through 
these community based avenues of communication is frowned apon.

@@

Working with QGIS is simple and intuitive as you are presented with a
modern and friendly graphical user interface (GUI) based on QT4. All
functionalities are clearly seperated (see Figure~\ref{fig:qgis10}).

@@

\textbf{Core Plugins} are maintained by the QGIS Development Team. They are
written in C++ or Python, are automatically part of every QGIS distribution
and can be loaded with the Plugin Manager. There are currently 17 core
plugins available, such as GRASS GIS integration (See
Figure~\ref{fig:grass-plugin}), Georeferencer, Mapserver Export, Shapefile to
PostGIS Import Tool, OGR Layer Converter, GPS Tools, Add Delimited Text Layer
or WFS support:

@@

\textbf{External Plugins} are all written in Python and devided into official
and user contributed plugins. The user can easily add those Plugins to QGIS
with the Python Plugin Installer (See Figure~\ref{fig:python-plugin}).

\begin{itemize}
\item \textbf{Official} external python plugins are stored in an official,
moderated repository at \url{http://pyqgis.org/repo/official} as part of the
official QGIS release and maintained by the individual author.
\item \textbf{User-Contributed} external python plugins are stored in an
unofficial repository at \url{http://pyqgis.org/repo/contributed} and contain
plugins that are not yet mature enough but on the way to the official
repository.
\end{itemize}

Beside these two respositories a few QGIS developers provide and maintain
their own repositories. They can be added to the repository list of the
Python Plugin Installer.

@@

As QGIS is open source software, it is possible to participate in development process and also to write new applications that use the libraries of the QGIS project. Development with QGIS can be done either in the existing classes of QGIS, as plugin extensions or in the form of custom applications that make use of the QGIS libraries. As a general note all code in QGIS is licensed under the GNU GPL. That means that for all three cases, published software must be distributed under the terms of the GPL too. QGIS 1.0 provides a stable API such that changes to developed plugins and applications will only be necessary for version 2 of QGIS.

\subsubsection{Development in the core classes of QGIS}
Changes in the already existing classes may be placed as patches at the bug tracker of the QGIS project (https://trac.osgeo.org/qgis/). The code maintainers of the QGIS project, each responsible for a certain part of the code base, regularly check the tracker.

\subsubsection{Development of extensions as C++ or Python plugins}
There is a plugin interface that allows extensions to access the running QGIS instance and to use and extend the objects in the core of QGIS. Plugins may be written in C++ or in Python. The QGIS documentation contains simple examples for both such that it is straightforward to start with plugin programming.

\subsubsection{Custom applications that use the QGIS libaries}
It is also possible to write new applications that provide its own user interface and use the QGIS core library for the GIS logic, data access and map rendering.

An example for this approach is the QGIS mapserver project that provides a WMS compatible mapserver on top of the QGIS core library. This software has not a graphical user interface. It is a FastCGI application that waits until called by a webserver. It parses the request parameters and uses QGIS to render a map into an offscreen buffer. The content is then returned as binary image back to the client.

Another context where this approach would make sense is to provide a mapping application for mobile devices. Applications for mobile devices usually need different user interfaces compared to applications for desktop computers and laptops. Still those applications could use the QGIS libraries as a GIS backend.

>>>>

The QGIS project is the work of a group of dedicated developers,
translators, documenters, release helpers, bug reporters, and promoters. Their 
contributions are mainly on a voluntary basis, except in a few cases where 
contributors are able to contribute to QGIS as part of their daily work. QGIS 
is managed by the Project Steering Committee (PSC), a five member committee
providing technical guidance, community liason, release management, and
financial/marketing activities. The work of the QGIS project process is
spread between numerous people who each have a specific area of
resposibility, and ad-hoc contributors.

These volunteers together with a large number of users make up the
world-wide QGIS comunity. Over time their efforts have resulted in a comprehensive,
valuable and useful code and documentation base which is free for everyone 
to use and improve apon.

@@

With community platforms such as our website, wiki, forums and blog the QGIS project
provides current news, release, usage, and development information. In most cases 
these community web sites permit user contributions after registering. 
The QGIS-user mailing, forum and Internet Relay Chat (IRC) provide a valuable interface 
with other users and for discussions of QGIS in general. In the spirit of open 
process and sharing knowledge, contacting developers directly instead of going through 
these community based avenues of communication is frowned apon.

@@

Working with QGIS is simple and intuitive as you are presented with a
modern and friendly graphical user interface (GUI) based on QT4. All
functions are clearly separated (see Figure~\ref{fig:qgis10}).

@@

\textbf{Core Plugins} are maintained by the QGIS Development Team. They are
written in C++ or Python, are automatically part of every QGIS distribution
and can be enabled with the Plugin Manager. There are currently 17 core
plugins available, including GRASS GIS integration (See
Figure~\ref{fig:grass-plugin}), Georeferencer, Mapserver Export, Shapefile to
PostGIS Import Tool, OGR Layer Converter, GPS Tools, Add Delimited Text Layer
and WFS support.

@@

\textbf{External Plugins} are all written in Python and divided into official
and user contributed plugins. The user can easily add those Plugins to QGIS
with the Python Plugin Installer (See Figure~\ref{fig:python-plugin}).

\begin{itemize}
\item \textbf{Official} external python plugins are stored in an official,
moderated repository at \url{http://pyqgis.org/repo/official} as part of the
official QGIS release and maintained by their respective author.
\item \textbf{User-Contributed} external python plugins are stored in an
unofficial repository at \url{http://pyqgis.org/repo/contributed} and contain
plugins that are not yet mature enough but on the way to the official
repository.
\end{itemize}

Beside these two respositories a number of QGIS developers provide and maintain
their own repositories. These can be added to the repository list of the
Python Plugin Installer.

@@

Since QGIS is open source software, it is possible to participate in the development
process and also to write new applications that use the libraries of the QGIS
project. Development with QGIS can be done either in the existing classes of
QGIS, as plugin extensions or in the form of custom applications that make use
of the QGIS libraries. As a general note all code in QGIS is licensed under the
GNU GPL. That means that for all three cases, published software must be
distributed under the terms of the GPL too. QGIS 1.0 provides a stable API which 
provides an assurance that plugins and applications developed against the 1.0 API 
will work against future releases in the 1.X release series.

\subsubsection{Development in the core classes of QGIS}
Changes to existing classes may be submitted as patches using QGIS Project bug
tracker (https://trac.osgeo.org/qgis/). The code maintainers of the QGIS
project, each responsible for a certain part of the code base, regularly check
the tracker.

\subsubsection{Development of extensions as C++ or Python plugins}
There is a plugin interface that allows extensions to access the running QGIS
instance and to use and extend the objects in the core of QGIS. Plugins may be
written in C++ or in Python. The QGIS documentation contains simple examples
for both programming languages making it straightforward to start with plugin 
programming.

\subsubsection{Custom applications that use the QGIS libaries}
It is also possible to write new applications that provide their own user
interface and use the QGIS core library for the GIS logic, data access and map
rendering.

An example using this approach is the QGIS mapserver project that provides a WMS
compatible mapserver on top of the QGIS core library. This software has no 
graphical user interface. It is a FastCGI application that waits until called
by a webserver. It parses the request parameters and uses QGIS to render a map
into an offscreen buffer. The content is then returned as binary image back to
the client.

Another context where this approach would make sense is to provide a mapping
application for mobile devices. Applications for mobile devices usually need
different user interfaces to desktop computers applications and
laptops. The QGIS libraries offer the potential to be used as a GIS backend for 
applications targeted to mobile devices.
